---
title: "Tracksuit Survey Allocation Optimisation"
author: "Cai Yahui"
date: "2026-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lpSolve)

set.seed(739)
```

# Problem Overview

Tracksuit runs monthly brand tracking surveys across multiple categories.

Each category requires approximately **200 qualified respondents per month**.

However, categories vary in:

- incidence rate (probability of qualifying)
- survey length

Additionally:

- the mean survey length must remain **below 480 seconds**
- the goal is to **minimise the number of respondents surveyed**

This report explores an algorithmic approach to allocating category qualifiers into survey structures.

# Load Data

```{r}
categories <- read_csv("../data/fake_category_data.csv")

categories <- categories %>%
  mutate(
    expected_time = incidence_rate * category_length_seconds,
    required_exposures = ceiling(200 / incidence_rate)
  )

head(categories)
```

We calculate the expected time contribution of each category:

$$expected\_time_i = incidence\_rate_i \times category\_length\_seconds_i$$

# Baseline Method - Greedy Packing

The baseline approach treats the problem as a **bin-packing optimisation** problem.

Categories are packed into survey bundles while ensuring:

$$Total\ expected\ time \leq 480\ seconds$$. 

```{r}
pack_greedy <- function(categories, max_time = 480){

  categories <- categories %>%
    arrange(desc(expected_time))

  surveys <- list()
  survey_time <- c()

  for(i in 1:nrow(categories)){

    cat <- categories[i,]
    placed <- FALSE

    for(j in seq_along(surveys)){

      if(survey_time[j] + cat$expected_time <= max_time){

        surveys[[j]] <- c(surveys[[j]], cat$category_id)
        survey_time[j] <- survey_time[j] + cat$expected_time
        placed <- TRUE
        break
      }
    }

    if(!placed){
      surveys[[length(surveys)+1]] <- c(cat$category_id)
      survey_time <- c(survey_time, cat$expected_time)
    }

  }

  surveys
}

greedy_design <- pack_greedy(categories)

greedy_design

```

# Improved Method – Demand Weighted Packing

The greedy algorithm does not consider how many respondents are required for each category.

Categories with **low incidence rates require more respondents** to reach the quota.

To address this, we introduce a priority score:

$$priority_i = required\_exposures_i \times expected\_time_i$$

This prioritises categories that both:
- require many respondents
- contribute significant survey time

```{r}
categories <- categories %>%
  mutate(priority = required_exposures * expected_time)

pack_weighted <- function(categories, max_time = 480){

  categories <- categories %>%
    arrange(desc(priority))

  surveys <- list()
  survey_time <- c()

  for(i in 1:nrow(categories)){

    cat <- categories[i,]
    placed <- FALSE

    for(j in seq_along(surveys)){

      if(survey_time[j] + cat$expected_time <= max_time){

        surveys[[j]] <- c(surveys[[j]], cat$category_id)
        survey_time[j] <- survey_time[j] + cat$expected_time
        placed <- TRUE
        break
      }
    }

    if(!placed){

      surveys[[length(surveys)+1]] <- c(cat$category_id)
      survey_time <- c(survey_time, cat$expected_time)

    }
  }

  surveys
}

weighted_design <- pack_weighted(categories)

weighted_design

```

# Linear Programming

While heuristic methods such as greedy packing are computationally efficient, they do not guarantee an optimal allocation.

To further improve the solution, we formulate the problem as an optimisation task using **Linear Programming (LP)**.

The objective is to minimise the total number of respondents required while ensuring that each category receives at least 200 qualified respondents.

The expected number of qualified respondents for a category is:

$$qualified_i = respondents_i \times incidence_rate_i$$

Therefore the minimum respondents required is:

$$respondents \geq 200 / incidence\_rate$$

This optimisation provides a theoretical lower bound for the number of respondents required.

```{r}
categories <- categories %>%
  mutate(
    required_respondents = ceiling(200 / incidence_rate)
  )
```

The LP problem can be expressed as:

Minimise:

total respondents

Subject to:

$$respondents \times incidence_rate \leq 200$$

```{r}
solve_lp <- function(categories){

  n <- nrow(categories)

  objective <- rep(1, n)

  constraint_matrix <- diag(categories$incidence_rate)

  constraint_dir <- rep(">=", n)

  constraint_rhs <- rep(200, n)

  solution <- lp(
    direction = "min",
    objective.in = objective,
    const.mat = constraint_matrix,
    const.dir = constraint_dir,
    const.rhs = constraint_rhs
  )

  solution$solution
}

lp_result <- solve_lp(categories)

lp_result
```

The LP solution estimates the minimum number of respondents required for each category to achieve 200 qualified responses in expectation.

However, because qualification is probabilistic, the LP result represents a **theoretical lower bound** rather than a guaranteed solution.

# Simulation Validation

Because qualification is probabilistic, we validate the policy using Monte Carlo simulation.

```{r}
run_simulation <- function(categories, survey_design, n_runs = 500){

  map_dfr(1:n_runs, function(run){

    qualified_counts <- rep(0, nrow(categories))
    names(qualified_counts) <- categories$category_id

    total_time <- 0
    respondents <- 0

    while(any(qualified_counts < 200)){

      respondents <- respondents + 1

      survey <- survey_design[[sample(1:length(survey_design),1)]]

      for(cat in survey){

        inc <- categories$incidence_rate[
          categories$category_id == cat
        ]

        length <- categories$category_length_seconds[
          categories$category_id == cat
        ]

        qualify <- rbinom(1,1,inc)

        if(qualify==1){

          qualified_counts[as.character(cat)] <-
            qualified_counts[as.character(cat)] + 1

          total_time <- total_time + length
        }

      }

    }

    tibble(
      respondents = respondents,
      avg_time = total_time/respondents
    )

  })
}

```

```{r}
greedy_results <- run_simulation(categories, greedy_design)

weighted_results <- run_simulation(categories, weighted_design)
```

## Simulation Using LP Baseline

```{r}
simulate_lp <- function(categories, n_runs = 500){

  map_dfr(1:n_runs, function(run){

    respondents <- ceiling(max(200/categories$incidence_rate))

    qualified <- rbinom(
      nrow(categories),
      respondents,
      categories$incidence_rate
    )

    tibble(
      respondents = respondents,
      success = all(qualified >= 200)
    )

  })
}

lp_sim <- simulate_lp(categories)

lp_sim

```


```{r}
comparison <- tibble(
  method = c("Greedy","Weighted","LP lower bound"),
  respondents = c(
    mean(greedy_results$respondents),
    mean(weighted_results$respondents),
    mean(lp_sim$respondents)
  )
)

comparison
```


# Visualisation

```{r}
ggplot(greedy_results, aes(respondents)) +
  geom_histogram(bins=30, fill="steelblue") +
  labs(
    title="Distribution of Respondents Required for Greedy Method",
    x="Respondents",
    y="Frequency"
  )

ggplot(weighted_results, aes(respondents)) +
  geom_histogram(bins=30, fill="steelblue") +
  labs(
    title="Distribution of Respondents Required for Weighted Method",
    x="Respondents",
    y="Frequency"
  )
```


```{r}
ggplot(comparison, aes(method, respondents, fill = method)) +
  geom_bar(stat="identity") +
  labs(
    title="Respondent Cost Comparison",
    x="Method",
    y="Respondents Required"
  ) +
  theme_minimal()
```

# Survey Length Validation

The simulation confirms that the expected interview length remains below the 480 second constraint.

```{r}
mean(greedy_results$avg_time)
mean(weighted_results$avg_time)
```


## Why Compare These Approaches?

Three approaches are evaluated:

Greedy packing  
Simple heuristic that prioritises categories with large expected survey time.

Demand-weighted packing  
Improves the heuristic by prioritising categories that require many respondents.

Linear Programming baseline  
Provides a theoretical lower bound for the number of respondents required.

Simulation results show that the demand-weighted heuristic approaches the LP lower bound while remaining simple and scalable.   


# Real-World Considerations

Several extensions could further improve this system:

1. **Incidence uncertainty**

Estimated incidence rates may differ from real values.

2. **Adaptive allocation**

Survey allocation could be dynamically adjusted during the month.

3. **Audience clustering**

Categories with similar audiences could potentially be grouped together to increase qualification efficiency. However, this approach risks introducing demographic bias and would require careful validation.

# Conclusion

This project demonstrates a scalable approach to survey allocation by combining:

- heuristic optimisation
- demand-aware prioritisation
- Linear Programming Optimisation
- Monte Carlo simulation

The improved allocation strategy reduces the number of respondents required while maintaining survey quality constraints.
While the LP solution provides an optimal theoretical benchmark, the demand-weighted heuristic offers a practical and scalable strategy for production deployment.

**Note** : Because qualification events are stochastic (Bernoulli trials), quota fulfilment becomes a stopping-time process. As a result, simulation is required to evaluate the probabilistic performance of each allocation strategy.








